%=====================================================================
%  references.tex — References (Rules 9.1–9.3)
%
%  STYLE  : APA (Latest). Scholars of the Faculty of Engineering and
%           CS&IT may instead follow their faculty-approved style (IEEE).
%  FORMAT : font size 11, SINGLE line spacing, space before each entry
%           (the singlespace + parskip below handle this automatically).
%  ORDER  : APA — list alphabetically by first author's surname.
%           IEEE — list in citation order and number [1], [2], ...
%  RULE   : every reference here must be cited in the text.
%
%  Journal, book, and book-chapter entries use DIFFERENT patterns.
%  One model of each is given below — copy the matching one.
%=====================================================================
\begin{singlespace}
\setlength{\parskip}{12pt}


% The {99} tells LaTeX to expect up to 99 references so it aligns the numbers properly
\begin{thebibliography}{99}
	
	% --- Reference 1 ---
	% \bibitem{keyword} is what you use to link the text to this reference.
% 1. Abedin & Mehrub (First citation in 3.1)
\bibitem{abedin2026}
M. M.-H.-Z. Abedin and T. Mehrub, ``Comprehensive Review on Deep Learning and Foundation Models in Agriculture: Progress, Benchmarks, and Open Challenges,'' 2026.

% 2. YOLOv11 13.5ms inference (Anonymous)
\bibitem{sharma2024}
A. Sharma, V. Kumar, and L. Longchamps, ``Comparative performance of YOLOv8, YOLOv9, YOLOv10, YOLOv11 and Faster R-CNN models for detection of multiple weed species,'' 2024.

% 3. Raspberry Pi 5-layer CNN (Anonymous)
\bibitem{razfar2022}
N. Razfar et al., ``Weed detection in soybean crops using custom lightweight deep learning models,'' 2022.

% 4. Ghose et al. 
\bibitem{ghose2025}
P. Ghose et al., ``YOLO-SAM AgriScan: A Unified Framework for Ripe Strawberry Detection and Segmentation with Few-Shot and Zero-Shot Learning,'' 2025.

% 5. 15-20 minutes annotation time (Anonymous)
\bibitem{gao2025a}
X. Gao, J. Gao, and W. A. Qureshi, ``Applications, Trends, and Challenges of Precision Weed Control Technologies Based on Deep Learning and Machine Vision,'' 2025.

% 6. Li et al. (Cross-continental crop classification)
\bibitem{li2025}
S. Li et al., ``A Class-Aware Unsupervised Domain Adaptation Framework for Cross-Continental Crop Classification with Sentinel-2 Time Series,'' 2025.

% 7. Deng et al., 2024 (Cross-season detection / DINO drops)
\bibitem{deng2024}
B. Deng, Y. Lu, and J. Xu, ``Weed database development: An updated survey of public weed datasets and cross-season weed detection adaptation,'' 2024.

% 8. Faster R-CNN plummet to 9.55% (Anonymous)
\bibitem{gao2024}
J. Gao et al., ``Cross-domain transfer learning for weed segmentation and mapping in precision farming using ground and UAV images,'' 2024.

% 9. Nadeem et al., 2026 (CCUDA-SS)
\bibitem{nadeem2026}
N. Nadeem, M. H. Asad, and A. Bais, ``CCUDA-SS: Cross-Crop Unsupervised Domain Adaptation for Semantic Segmentation,'' 2026.

% 10. Ilyas et al., 2023 (Overcoming field variability)
\bibitem{ilyas2023}
T. Ilyas et al., ``Overcoming field variability: unsupervised domain adaptation for enhanced crop-weed recognition in diverse farmlands,'' 2023.

% 11. Deng et al., 2025 (WeedTeacher)
\bibitem{deng2025}
B. Deng, Y. Lu, and D. Brainard, ``Semi-Supervised Weed Detection in Vegetable Fields: In-domain and Cross-domain Experiments,'' 2025.

% 12. ViT parameters reduced 30% (Anonymous)
\bibitem{ahmad2024}
S. Ahmad et al., ``AI-Enabled Vision Transformer for Automated Weed Detection: Advancing Innovation in Agriculture,'' 2024.

% 13. Islam et al., 2025 (WeedSwin)
\bibitem{islam2025}
T. Islam et al., ``WeedSwin hierarchical vision transformer with SAM-2 for multi-stage weed detection and classification,'' 2025.

% 14. Grounded YOLOv10-SAM2.1 (Anonymous)
\bibitem{verma2026}
P. Verma and A. Paul, ``Toward Grounded YOLO-SAM: Unified Detection–Segmentation Framework for Agricultural Intelligence,'' 2026.

% 15. Chong et al., 2024 (Zero-Shot Semantic Segmentation)
\bibitem{chong2025}
Y. L. Chong et al., ``Zero-Shot Semantic Segmentation for Robots in Agriculture,'' 2025.

% 16. Soniya et al., 2024 (When YOLO meets SAM)
\bibitem{soniya2025}
Soniya, N. Ambadipudi, and A. Narayan, ``When YOLO Meets SAM: Data-Efficient Weed Density Estimation,'' 2025.

% 17. Reliance on standard RGB cameras (Anonymous)
\bibitem{hamrani2025}
A. Hamrani et al., ``AI and Robotics in Agriculture: A Systematic and Quantitative Review of Research Trends (2015–2025),'' 2025.

% 18. WeedsGalore dataset 63.08% (Anonymous)
\bibitem{galymzhankyzy2025}
Z. Galymzhankyzy and E. Martinson, ``Lightweight Multispectral Crop-Weed Segmentation for Precision Agriculture,'' 2025.

% 19. DeepLabv3+ massive 41.2 million params (Anonymous)
\bibitem{yenna2026}
G. R. Yenna et al., ``HybridViT-CAB: a vision transformer and convolutional attention network for precision weed detection in agricultural systems,'' 2026.

% 20. S3-Net framework (Song et al. / Shi et al.)
\bibitem{shi2026}
L. Shi et al., ``An AI-Driven Dual-Spectral Vision–Language Sensing Framework for Intelligent Agricultural Phenotyping,'' 2026.

% --- REMAINING UNUSED PAPERS (Sorted below so they don't mess up your sequence) ---

\bibitem{dosovitskiy2020}
A. Dosovitskiy et al., ``An Image is Worth 16x16 Words: Transformers for Image Recognition at Scale,'' 2020.

\bibitem{ogidi2023}
F. C. Ogidi, M. G. Eramian, and I. Stavness, ``Benchmarking Self-Supervised Contrastive Learning Methods for Image-Based Plant Phenotyping,'' 2023.

\bibitem{mo2025}
H. Mo et al., ``CDGLYOLOv11Lite: A Lightweight Field Maize–Seedling Weed Detector Based on Cross-Domain Shared Attention and a Conditional Gated Linear Unit,'' 2025.

\bibitem{luca2025a}
C. Luca, ``Comparison of Vision Transformers and CNNs in the Context of Agricultural Weed Detection,'' 2025.

\bibitem{tzouras2023}
V. Tzouras, L. Nalpantidis, and R. Güldenring, ``From Web Data to Real Fields: Low-Cost Unsupervised Domain Adaptation for Agricultural Robots,'' 2023.

\bibitem{magistri2023}
F. Magistri et al., ``From one field to another—Unsupervised domain adaptation for semantic segmentation in agricultural robotics,'' 2023.

\bibitem{modak2024}
S. Modak and A. Stein, ``Generative AI-based Pipeline Architecture for Increasing Training Efficiency in Intelligent Weed Control Systems,'' 2024.

\bibitem{li2024}
J. Li et al., ``Performance evaluation of semi-supervised learning frameworks for multi-class weed detection,'' 2024.

\bibitem{simsek2026}
F. F. Şimşek and M. Altay, ``Phenology aware agricultural boundary extraction using segment anything model and planet scope imagery (zero shot learning approach),'' 2026.

\bibitem{luca2025b}
C. Luca, ``Transfer Learning and Fine-Tuning Vision Transformers for Multi-Crop Weed Classification,'' 2025.

\bibitem{huang2024}
Y. Huang and A. Bais, ``Unsupervised Domain Adaptation for Weed Segmentation Using Greedy Pseudo-labelling,'' 2024.

\bibitem{pei2022}
H. Pei et al., ``Weed Detection in Maize Fields by UAV Images Based on Crop Row Preprocessing and Improved YOLOv4,'' 2022.

	
% ---------- APA: Journal article ----------
% Author, A. A., & Author, B. B. (Year). Title of article. Journal
% Name, Volume(Issue), pages.


% ---------- APA: Book ----------
% Author, A. A. (Year). Title of the book (Edition). Publisher.


% ---------- APA: Book chapter ----------
% Author, A. A. (Year). Title of chapter. In E. Editor (Ed.), Title of
% book (pp. xx–xx). Publisher.


% ---------- IEEE: Conference paper (Engineering / CS&IT) ----------
% [n] A. Author, B. Author, and C. Author, ``Paper title,'' in Proc.
%     Conf. Name (ABBR), Year, pp. xx–xx.
% [1] Z. Wang, L. Zheng, Y. Liu, Y. Li, and S. Wang, ``Towards real-time
%     multi-object tracking,'' in \textit{Proc. ECCV}, 2020, pp. 107–122.
\end{thebibliography}
\end{singlespace}
